\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2023-07-04 cubble.Rmd
use \code{} markup

-   a spatial cubble with class `c("spatial_cubble_df", "cubble_df")`
-   a temporal cubble with class `c("temporal_cubble_df", "cubble_df")`

In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised `ts` column. For example, the spatial cubble object, \code{cb_spatial} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database [@menne2012overview]. In this case, the spatial cubble is convenient for wrangling the spatial variables:

@@

Matching temporal data across different locations from different data sources could be done by initially identifying the corresponding spatial locations between the two data sets. Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to matched the selected time series across locations. In cubble, locations from two datasets can be matched using the function `match_spatial()`. The functi on calculates the distance matrix of the locations between the two data sets and returns  groups (`spatial_n_group`) with the smallest distances. For a given group, it is possible can include more locations with the argument `spatial_n_each` (default to 1 for one-on-one matching).

For the temporal matching a similarity score between the time series of spatially matched pairs is computed using the function `match_temporal()`. The similarity score is computed by a matching function which can be customized to any desired time series feature. The function `match_temporal()` takes as argument two time series in the form of a list and returns a single numerical value. By default, cubble uses a simple peak matching algorithm (`match_peak`) to count the number of peaks in two time series that fall within a specified time window.

The temporal matching requires two identifiers: one for separating each spatially matched group: `match_id` and one for separating the two data sources: `data_id`. Matching between different variables can be specified using the `temporal_by` argument, similar to the `by` syntax from dplyr's `*_join`.

>>>>

-   a spatial cubble with class \code{c("spatial_cubble_df", "cubble_df")}
-   a temporal cubble with class \code{c("temporal_cubble_df", "cubble_df")}

In a spatial cubble object, spatial variables are organised as columns and temporal variables are nested within a specialised \code{ts} column. For example, the spatial cubble object, \code{cb_spatial} printed below, contains weather records of three airport stations from the Global Historical Climatology Network Daily (GHCND) database [@menne2012overview]. In this case, the spatial cubble is convenient for wrangling the spatial variables:

@@

Matching temporal data across different locations from different data sources could be done by initially identifying the corresponding spatial locations between the two data sets. Subsequently, a set of temporal features can be calculated for the series at the selected locations that can be used to matched the selected time series across locations. In cubble, locations from two datasets can be matched using the function \code{match_spatial()}. The functi on calculates the distance matrix of the locations between the two data sets and returns  groups (\code{spatial_n_group}) with the smallest distances. For a given group, it is possible can include more locations with the argument \code{spatial_n_each} (default to 1 for one-on-one matching).

For the temporal matching a similarity score between the time series of spatially matched pairs is computed using the function \code{match_temporal()}. The similarity score is computed by a matching function which can be customized to any desired time series feature. The function \code{match_temporal()} takes as argument two time series in the form of a list and returns a single numerical value. By default, cubble uses a simple peak matching algorithm (\code{match_peak}) to count the number of peaks in two time series that fall within a specified time window.

The temporal matching requires two identifiers: one for separating each spatially matched group: \code{match_id} and one for separating the two data sources: \code{data_id}. Matching between different variables can be specified using the \code{temporal_by} argument, similar to the \code{by} syntax from dplyr's \code{*_join}.
